Efficiently summarising relationships in large samples: a general duality between statistics of genealogies and genomes

### 2019-09-06 treestats_paper.tex
why trees differ

Before defining the different classes of statistic,
it will help to establish a little more notation.
A tree sequence efficiently describes how a set of $n$ sampled chromosomes
are related to each other along a (linear) genome of length $L$ \citep{kelleher2016efficient,kelleher2018efficient}.
% These trees are highly correlated,
% a fact which allows them to be stored and processed very efficiently,
% as described in \citet{kelleher2018efficient}.
From the tree sequence can be extracted a sequence of trees,
$\treeseq = (T_1, T_2, \ldots, T_{n(\treeseq)})$
and a sequence of breakpoints $0 = a_0 < a_1 < \cdots < a_{n(T)} = L$,
where $T_k$ describes the genealogical relationships of the samples
over the segment of genome betwen $a_{k-1}$ (inclusive) and $a_k$ (exclusive).
We say that tree $T_k$ \emph{covers} the (half-open) segment $[a_{k-1}, a_k)$,
and call the length of this segment its \emph{span}, denoted $L_k = a_k - a_{k-1}$.
We refer to the branches in each tree using the most recent node,
so for instance the branch describing a relationship where $u$ is the child of $v$
is associated with $u$.
The \emph{length} of this branch is the difference in birth times of $v$ and $u$,
and is denoted $t_T(u)$ (dropping $T$ if it is clear from context).
Note that although the same parent-child relationship may exist across many adjacent trees
(this is called an ``edge''),
rearrangements of genealogical relationships
can cause the precise set of samples that lie below that edge to differ across
trees.

>>>>

Before defining the different classes of statistic,
it will help to establish a little more notation.
A tree sequence efficiently describes how a set of $n$ sampled chromosomes
are related to each other along a (linear) genome of length $L$ \citep{kelleher2016efficient,kelleher2018efficient}.
% These trees are highly correlated,
% a fact which allows them to be stored and processed very efficiently,
% as described in \citet{kelleher2018efficient}.
From the tree sequence can be extracted a sequence of trees,
$\treeseq = (T_1, T_2, \ldots, T_{n(\treeseq)})$
and a sequence of breakpoints $0 = a_0 < a_1 < \cdots < a_{n(T)} = L$,
where $T_k$ describes the genealogical relationships of the samples
over the segment of genome betwen $a_{k-1}$ (inclusive) and $a_k$ (exclusive).
We say that tree $T_k$ \emph{covers} the (half-open) segment $[a_{k-1}, a_k)$,
and call the length of this segment its \emph{span}, denoted $L_k = a_k - a_{k-1}$.
We refer to the branches in each tree using the most recent node,
so for instance the branch describing a relationship where $u$ is the child of $v$
is associated with $u$.
The \emph{length} of this branch is the difference in birth times of $v$ and $u$,
and is denoted $t_T(u)$ (dropping $T$ if it is clear from context).
Note that although the same parent-child relationship may exist across many adjacent trees
(this is called an ``edge''),
rearrangements of genealogical relationships
can cause the precise set of samples that lie below that edge to differ across
trees due to recombination.
